Paper

### 2024-12-14 paper/index.qmd
odes work

>>>>

```{r}
single_outbreak_sim_mod <- single_outbreak_sim %>% 
  mutate(prop_infected = `current-colonised` / `current-inpatients`,
         current_not_infected = `current-inpatients` - `current-colonised`)
single_outbreak_sim_mod %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  ggplot(aes(x = `[step]`, y = value, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment")

# single_outbreak_sum_stat <- single_outbreak_sim_mod %>% 
#   group_by(`[step]`) %>%
#   summarise(m = mean(prop_infected))
```

```{r}
# Parameters and time-dependent functions
parameters <- list(
  # p = 1 - unique(single_outbreak_sim$`community-colonisation-rate`),      # Proportion of non-infected admissions
  A = 50        # Total admission/discharge rate
)
#

single_seeds <- single_outbreak_sim$`random-seed`
single_outbreak_one_seed <- single_outbreak_sim_mod %>% 
  filter(`random-seed` == single_seeds[1])

target_baseline_rate <- single_outbreak_one_seed %>%
  filter(`[step]` > 30 & `[step]` < 200) %>%
  summarise(prop_infected = mean(prop_infected)) %>% 
  pull(prop_infected)
target_outbreak_rate <- single_outbreak_one_seed %>%
  filter(`[step]` > 1100 & `[step]` < 1200) %>%
  summarise(prop_infected = mean(prop_infected)) %>% 
  pull(prop_infected)

baselinerate <- seq(from=0.005, to=0.1, length.out = 100)
outbreakrate <- seq(0.005, 0.1, length.out = 100)
baselinecolonisation <- seq(0.9, 0.99, length.out = 100)
outbreakcolonisation <- seq(0.9, 0.99, length.out = 100)
rate_grid <- expand.grid(b_base = baselinerate,
                         b_out = outbreakrate,
                         c_base = baselinecolonisation,
                         c_out = outbreakcolonisation)
rate_grid <- rate_grid %>% 
  filter(b_base < b_out & c_base > c_out) %>% 
  slice_sample(n = 100)

# Time-varying beta function (example: sinusoidal variation over time)
beta_function <- function(t, baseline_rate, outbreak_rate) {
  if (t > unique(single_outbreak_sim$`outbreak-start`) &
      t < unique(single_outbreak_sim$`outbreak-end`)) {
    return (outbreak_rate)
  }
  return (baseline_rate)
}

# Initial conditions
initial_state <- c(N = 600, I = 17)  # Initial populations of non-infected and infected patients
time <- seq(0, run_to, by = 1)         # Time steps
dt <- 1                             # Time step size

plan(multisession)
ode_sims <- furrr::future_pmap_dfr(rate_grid,
        \(b_base, b_out, c_base, c_out) {
          N <- numeric(length(time))
          I <- numeric(length(time))
          N[1] <- initial_state["N"]
          I[1] <- initial_state["I"]

          # Euler method for numerical integration with time-varying parameters
          for (t in 1:(length(time) - 1)) {
            # Current beta value
            beta_t <- beta_function(time[t], b_base, b_out)
            p_t <- beta_function(time[t], c_base, c_out)

            # Total population at current time
            total <- N[t] + I[t]

            # Proportions of non-infected and infected
            prop_N <- ifelse(total > 0, N[t] / total, 0)
            prop_I <- ifelse(total > 0, I[t] / total, 0)

            # Calculate derivatives
            dN <- p_t * parameters$A - parameters$A * prop_N - beta_t * N[t] * prop_I
            dI <- (1 - p_t) * parameters$A - parameters$A * prop_I + beta_t * N[t] * prop_I

            # Update populations using Euler's method
            N[t + 1] <- N[t] + dN * dt
            I[t + 1] <- I[t] + dI * dt
          }
          data.frame(baseline = mean(I[31:200] / ( N[31:200] + I[31:200] )),
                     outbreak = mean(I[1100:1200] / ( N[1100:1200] + I[1100:1200] )))
        })

abc(c(target_baseline_rate, target_outbreak_rate), rate_grid, ode_sims[c("baseline", "outbreak")], tol=0.5, method='rejection') %>% 
  summary()

baselinerate <- 0.0236
outbreakrate <- 0.0482

# baselinecolonisation <- 1 - unique(single_outbreak_sim$`community-colonisation-rate`)
# outbreakcolonisation <- 1 - unique(single_outbreak_sim$`o-community-colonisation-rate`)
baselinecolonisation <- 0.9721
outbreakcolonisation <- 0.9403

# Initialize vectors to store results
N <- numeric(length(time))
I <- numeric(length(time))
N[1] <- initial_state["N"]
I[1] <- initial_state["I"]

# Euler method for numerical integration with time-varying parameters
for (t in 1:(length(time) - 1)) {
  # Current beta value
  beta_t <- beta_function(time[t], baselinerate, outbreakrate)
  p_t <- beta_function(time[t], baselinecolonisation, outbreakcolonisation)

  # Total population at current time
  total <- N[t] + I[t]

  # Proportions of non-infected and infected
  prop_N <- ifelse(total > 0, N[t] / total, 0)
  prop_I <- ifelse(total > 0, I[t] / total, 0)

  # Calculate derivatives
  dN <- p_t * parameters$A - parameters$A * prop_N - beta_t * N[t] * prop_I
  dI <- (1 - p_t) * parameters$A - parameters$A * prop_I + beta_t * N[t] * prop_I

  # Update populations using Euler's method
  N[t + 1] <- N[t] + dN * dt
  I[t + 1] <- I[t] + dI * dt
}

# Combine results into a data frame for plotting
output_df <- data.frame(
  time = time,
  N = N,
  I = I
)

output_df <- filter(output_df, time > 31)

# Plot the results
library(ggplot2)
output_long <- reshape2::melt(output_df, id = "time")
ggplot(output_long, aes(x = time, y = value, color = variable)) +
  geom_line(size = 1) +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  scale_color_manual(values = c("blue", "red"),
                     labels = c("Non-Infected (N)", "Infected (I)")) +
  theme_minimal()

single_outbreak_one_seed %>% 
  pivot_longer(cols = c(`current-colonised`, `current_not_infected`), names_to = "compartment", values_to = "value") %>%
  group_by(compartment, `[step]`) %>% 
  summarise(m = mean(value)) %>% 
  ggplot(aes(x = `[step]`, y = m, color = compartment)) +
  geom_line() +
  labs(title = "Dynamics of Non-Infected and Infected Patients (Time-Varying Beta)",
       x = "Time",
       y = "Number of Patients",
       color = "Compartment") +
  geom_line(data = output_df, aes(x = time, y = N), color = "blue") + 
  geom_line(data = output_df, aes(x = time, y = I), color = "red")
```
